\documentclass[parskip=half, american]{scrartcl}
\usepackage{booktabs}
\usepackage[margin=2.5cm]{geometry}
\usepackage[dvipsnames]{xcolor}
\usepackage[colorlinks=true,
			citecolor=blue,
			urlcolor=blue,
			linkcolor=blue,
			pdfborder={0 0 0}]
	{hyperref}
	
\newcommand{\AR}[1]
	{\color{PineGreen}AR: #1\color{black} \par }
	
\newcommand{\journal}{\emph{Fire}}


\title{\Large Remote sensing in rangeland fire ecology: Comparing imagery to measured fire behavior, and burn severity across prescribed burns and wildfires  }
\subtitle{Response to review}
\author{ }
\date{}

\pagenumbering{gobble}
\raggedright 

\begin{document}

\maketitle

\vspace{-2cm} 

\AR{Author Responses (AR) appear below in this green text.}

\section*{Academic editor}

We received comments from 3 reviewers, all of whom provided good suggestions to improve the manuscript. Please address all review comments provided and respond with a detailed point-by-point response. 
Please note that one of the reviewers encouraged the use of a combined results and discussion section. The guide to authors lists the following as the recommended format of a manuscript: abstract, keywords, introduction, materials and methods, results, discussion, and conclusions (optional) sections. From this perspective, I encourage you to keep results and discussions separate. 
However, the main issue raised by the reviewer regarding more discussion around mechanisms is valid and needs to be discussed (reviewer 2 also raises this as an issue). 
It is also worth noting that the reviewers all commented on additional analysis. 

\AR{I appreciate the opportunity to consider these thoughtful reviews and revise accordingly. Major changes to this version include: 
	
\begin{enumerate}
	\item \emph{Higher-quality data with stronger trends.} 
	While not identified by reviewers, since submission I learned of an important step in scene normalization that corrects timeseries imagery for potentially confounding sources of error among scenes. 
	I completely re-processed the remotely-sensed data accordingly and worked in additional statistical tests following reviewer comments. 
	No trends in the data changed and thus nothing in the interpretation and conclusions have been affected, although a couple statistical results were more robust and thus provide better discussion points. 
	\item \emph{Summary tables.}
	Two reviewers requested data summaries and these have been added to the Appendix.
	\item \emph{More discussion of mechanisms.}
	Several paragraphs have been added to the Discussion and other relevant ones moved up to ensure a thorough discussion of the trends observed and the processes underlying them prior to the broader discussion of management implications. 
\end{enumerate}



} 


\end{document}
