\documentclass[parskip=half, american]{scrartcl}
\usepackage{booktabs}
\usepackage[margin=2.5cm]{geometry}
\usepackage[dvipsnames]{xcolor}
\usepackage[colorlinks=true,
			citecolor=blue,
			urlcolor=blue,
			linkcolor=blue,
			pdfborder={0 0 0}]
	{hyperref}
	
\newcommand{\AR}[1]
	{\color{PineGreen}AR: #1\color{black} \par }
	
\newcommand{\journal}{\emph{Fire}}


\title{\Large Remote sensing in rangeland fire ecology: Comparing imagery to measured fire behavior, and burn severity across prescribed burns and wildfires  }
\subtitle{Response to review}
\author{ }
\date{}

\pagenumbering{gobble}
\raggedright 

\begin{document}

\maketitle

\vspace{-2cm} 

\AR{Author Responses (AR) appear below in this green text.}

\section*{Reviewer 1}

Just like last time, I still think this paper is worth publishing after major revision even though authors did not revise. 
To help you see exactly what needs fixing, I reorganized my comments as below.

\AR{I appreciate the continued support for publication of this paper.}

The assessment results of burn severity based on remote sensing data need to be verified using CBI-field data. 
However, authors did not mention this point. I believe this should be discussed in either Section 1.2 or the Discussion section.

\AR{The absence of ground-truthed data for fires that occurred several years before has been explicitly acknowledged in the Discussion section.} 
	
In Section 2.3, only basic methods (i.e., Linear regression models) were employed. Some spatial statistical approaches could enhance the depth of this study.

\AR{As there are no spatial components to the hypotheses, no spatial analyses were conducted.
	Linear regression is a perfectly acceptable method of analysis used at even the highest levels of research when linear patterns are hypothesized. 
	Certainly, mixed-effect models and joint tests within interaction terms exceed ``basic'' linear regression.}
	
I suggest combining the Results and Discussion sections into a single section (``Results and Discussion'').

\AR{The AE indicated that this change should not be made.}
	
In Results and Discussion sections, you need to use more words to explain results from mechanism view. 

\AR{Several additional paragraphs have been added to the Discussion to elaborate on underlying mechanisms. }




\end{document}
