\documentclass[parskip=half, 
			   fontsize=12pt,
			   paper=a4]				
{scrartcl}
\input{SetupOptions} % contains scrletter

\newcommand{\journal}{\emph{Fire}}


\vspace{-5em} 

\begin{document}
	\begin{letter}{Editorial staff\\
		\journal }
\setlength{\parindent}{12pt}

\opening{Dear Sir or Madam,}  
		
I have attached a revised draft for resubmission to \journal. 
I am excited to leverage the combination of unique datasets to address persistent questions about wildland fire management and monitoring: How do remotely-sensed burn severity metrics relate to terrestrial measurements of fire behavior? 
And how transferable might research from prescribed burns be to areas burned by wildfire? 
This analysis provides encouraging results that will be of interest to wildland fire managers in rangelands worldwide\textemdash ecosystems frequently neglected by the forest-heavy fire ecology community. 

This paper is not under consideration for publication elsewhere. 
As the sole author, I confirm my approval for submission. 

Please note that this submission follows e-mail correspondence with Shirley Yang and Riva Song from 20 June 2025 regarding a full APC waiver for this manuscript. 

A response to the reviews from the initial submission are included in this submission. 

I look forward to working with the editors and reviewers of \journal~to ensure the manuscript meets the standards of the journal. 

\vspace{-3em} 
\closing{} %eg. Regards
%\cc{another dude}
\end{letter}



\end{document}
