---
title: 'Remote sensing in rangeland fire ecology: Comparing imagery to measured fire behavior, and burn severity across prescribed burns and wildfires'
subtitle:  'Figures and results'
author: "DAM"
date: "Last compiled: `r Sys.Date()`"
output:
  html_document:
    theme: united
    highlight: tango
    toc: true
    toc_float: true
---
  
```{r setup, echo=FALSE, message = FALSE}
knitr::opts_chunk$set(echo = FALSE, warning= FALSE,
                      message= FALSE, eval = TRUE, 
                      fig.path = "./Revision2/", dev = c('png', 'pdf')) 
pacman::p_load(tidyverse, sf, ggpattern, pander, readxl, grid, gridExtra)
pacman::p_load_gh('devanmcg/wesanderson')
```


```{r load_data, eval = TRUE}
# Mapping 
load('../../albersEAC.Rdata')
 rx_sf <- read_sf( '../data/OriginalRxBurnPerims.gpkg', 'RxLocations')
 wf_sf <- read_sf('../data/SeverityComparison.gpkg', 'GreatPlainsModifiedWF')
 ngp <- read_sf('../data/MapStuff.gpkg', 'ngp')
 ngp_counties <- read_sf('../data/MapStuff.gpkg', 'ngp_counties')
 usa <- read_sf('../data/MapStuff.gpkg', 'usa')
 gp <- read_sf('../data/MapStuff.gpkg', 'gp')
 NGP <- read_sf('../data/MapStuff.gpkg', 'NGP_states') # L3 for MT, ND, SD
 NGP_fires <- read_sf('../data/MapStuff.gpkg', 'NGP_fires')
 load('../data/ppt_dat.Rdata')
 

# Data 
  load('../data/RxPointSeverity.Rdata')
  load('../data/BurnModels.Rdata')
  load('../data/BurnSeverityData.Rdata')
# Aesthetics
  blues <- c(wes_palette('Zissou1')[2], wes_palette('Zissou1')[1])
```

# Overview 

## Objectives 

* Determine the relationship between *in agris* fire behavior measurements and remotely-sensed burn severity
* Compare burn severity across planned (Rx fire) and unplanned (wildfire) events
* Assess utility of fuelbed NDVI as predictor of fire behavior and burn severity

## Hypotheses

1. Positive linear relationship between *in agris* fire behavior and RS burn severity (higher intensity removes more aboveground plant material)
2. No difference in burn severity among wildfires and prescribed fires (Rx fires demonstrate a wide range of burn severity that likely overlaps that of wildfires)
3. Negative relationship between burn severity and NDVI (intensity declines with fine fuel moisture, indicated by high NDVI)

# Maps 

```{r study_map}
 locations <- 
   bind_rows( 
    wf_sf %>% 
      filter(FireCode != "HayCreek_2018") %>%
     st_transform(albersEAC) %>% 
      st_centroid() %>%
      mutate(type = 'Wildfire',
             L3 = case_when(
               FireCode %in% c('Swather_2021','Hwy31101STWest_2022', 'CB00121_2021', 
                               'CoalSeamWest_2021', '1806HunkpapaCreek_2021') ~
                 'Northwestern Glaciated Plains', 
               TRUE ~ L3), 
      zone = ifelse(L3 == "Northwestern Great Plains", 'Western', 'Eastern'), 
      zone = fct_rev(zone)) %>%
      select(type, zone), 
    rx_sf %>%
      mutate(type = 'Prescribed fire', 
             zone = ifelse(location == "cgrec", 'Eastern', 'Western')) %>%
      select(type, zone) %>%
     st_transform(albersEAC) )
  
 states <-  
   ngp %>%
   mutate(Area = st_area(.)) %>%
   group_by(state) %>%
   summarize(Area = sum(Area))  

reg_gg <- 
  ppt_dat$ppt_grd %>%
    st_intersection(ngp, .) %>%
    filter(st_geometry_type(., by_geometry = TRUE) %in% c('MULTIPOLYGON', 'POLYGON'))  %>%
    mutate(anomaly = case_when(
      L3 == "Northwestern Glaciated Plains" ~ ppt - 447,  
      L3 == "Northwestern Great Plains" ~ ppt - 431 )) %>%
    ggplot() + theme_void() +
    geom_sf(aes(fill = anomaly), color = 'NA') +
    geom_sf_pattern(data =  ppt_dat$anom_map %>% filter(anom_cat == 'beyond'), 
                    aes(pattern = anom_cat), 
                    pattern_color = NA,
                    pattern_fill = "grey80",
                    pattern_angle = 45,
                    pattern_alpha = 0.5,
                    pattern_density = 0.5,
                    pattern_spacing = 0.025,
                    pattern_key_scale_factor = 1,, 
                    show.legend = F) +
    geom_sf(data = ngp_counties %>% 
  st_intersection(states %>% st_transform(st_crs(ngp_counties))), 
            fill = NA, color = 'grey70') + 
    geom_sf(data = states, fill = NA, color = 'black') + 
    geom_sf_text(data = states %>%
                   st_centroid(), 
                 aes(label = state),
                 fontface = 'bold',
                  color = 'black') + 
    geom_sf_label(data = locations %>% filter(type != 'Wildfire'), 
            aes(color = zone),
            label = expression(R[X]), 
            label.padding = unit(0.1, "lines"),
            label.r = unit(0.5, "lines"),
            label.size = 0.15,
            size = 5) +
    geom_sf(data = locations %>% filter(type == 'Wildfire'), 
            aes(shape = zone, color = zone), 
             stroke = 2, size = 3) +
    # geom_sf(data = locations, aes(shape = type, color = type), 
    #         fill = 'lightgreen', 
    #         stroke = 2, size = 3) +
    scale_fill_gradient2("Precipitation\nanomaly (mm)", 
                         low = wes_palette("Zissou1")[4],
                         mid = 'white',
                         high = wes_palette("Zissou1")[1], 
                         na.value = "grey70") +
    scale_pattern_manual(values = c(beyond = "circle", within = "none")) + 
    scale_shape_manual('Zone', values = c(4,3)) + 
    scale_color_manual('Fire type', values = c('darkgreen', 'darkred'), guide = 'none')
  
us_gg <- 
  ggplot() + theme_void() + 
 # geom_sf(data = usa %>% st_buffer(10000), color = NA, fill = 'white')  + 
  geom_sf(data = usa, color = 'white', fill = 'grey60')  + 
  geom_sf(data = gp, fill = 'white', 
          color =  'black', 
          alpha = 0.5, size = 2) +
  geom_sf(data = ngp, 
          fill =  wes_palette('Zissou1')[5],  
          color = NA, 
          show.legend = F) +
  scale_fill_manual('EPA Level III\necoregion', 
                    values = c("#74A089","#E1BD6D")) 
  
  
grid.newpage() 
main <- viewport(width = 1, height = 0.95, x = 0.5, y = 0.5) 
inset <- viewport(width = 0.3, height = 0.4, x = 0.725, y = 0.7)
print(reg_gg, vp = main)
print(us_gg, vp = inset)  
```  

```{r region_map}
reg_gg <-
ppt_dat$ngp_ppt %>%
  mutate(NA_L3NAME = str_replace(NA_L3NAME, ' ', '\n')) %>%
ggplot() + theme_void() +
  geom_sf(aes(fill = NA_L3NAME, alpha = ppt), 
          color = 'NA') +
  geom_sf(data = NGP, 
          fill = NA) + 
   geom_sf_label(data = locations %>% 
                  filter(type != 'Wildfire'), 
                label = expression(R[X]), 
                label.padding = unit(0.1, "lines"),
                label.r = unit(0.5, "lines"),
                label.size = 0.15,
                size = 4) +
  geom_sf(data = NGP_fires, 
          pch = 1, alpha = 0.5) + 
  geom_sf_label(data = NGP %>%
                  group_by(STATE_NAME) %>%
                  summarize(area = sum(Shape_Area)) , 
                aes(label = state.abb[match(STATE_NAME,state.name)]), 
                fontface='bold') + 
  scale_fill_manual('EPA Level III\necoregion', 
                    values = c("#74A089","#E1BD6D")) +
  scale_alpha_continuous("Precipitation\nmean annual (mm)", 
                         range = c(0.1, 1)) + 
  theme(legend.position = 'right', 
        legend.key.height = unit(1, 'cm')) + 
  guides(alpha = guide_legend(order = 2), fill = guide_legend(order = 1) ) 


us_gg <- 
  ggplot() + theme_void() + 
  geom_sf(data = usa %>% st_buffer(10000), color = NA, fill = 'white')  + 
  geom_sf(data = usa, color = 'white', fill = 'grey60')  + 
  geom_sf(data = gp, fill = 'white', 
          color =  'black', 
          alpha = 0.5, size = 2) +
  geom_sf(data = NGP, 
          fill =  wes_palette('Zissou1')[5],  
          color = NA, 
          show.legend = F) +
  scale_fill_manual('EPA Level III\necoregion', 
                    values = c("#74A089","#E1BD6D")) 


grid.newpage() 
main <- viewport(width = 1, height = 0.95, x = 0.5, y = 0.5) 
inset <- viewport(width = 0.3, height = 0.4, x = 0.25, y = 0.25)
print(reg_gg, vp = main)
print(us_gg, vp = inset)
```

```{r}
NGP_fires %>% 
  as_tibble() %>%
  group_by(STATE_NAME, NA_L3NAME) %>% 
  summarize(Fires = n()) %>%
  pivot_wider(names_from = NA_L3NAME, 
              values_from = Fires) %>%
   mutate(Total = `Northwestern Great Plains` + `Northwestern Glaciated Plains`) %>%
  #xtable::xtable() 
  pander::pander("Fire incidents within each LIII ecoregion by state.")
```

# Fire behavior and burn severity

## Burn severity as response

```{r dNBR_v_FB, fig.width = 8, fig.height = 3.5, fig.cap = " "}
  RxPointSeverity %>%
  filter(between(dNBR, 0, 0.6)) %>%
    mutate(ros = ifelse(ros >= 13, NA, ros)) %>%
    pivot_longer(names_to = 'behavior', 
                 values_to = 'value', 
                 cols = c(MaxC:ros)) %>%
    mutate(behavior = fct_relevel(behavior, c('MaxC', 'SoilMaxC', 'ros')), 
           Behavior = recode(behavior, 
                        ros = 'Spread rate (m/min)', 
                        SoilMaxC = 'Soil surface temp (°C)',
                        MaxC = 'Flame temp (°C)') )   %>%
 
    ggplot(aes(x = value, y = dNBR)) + theme_bw(16) +
    geom_smooth(data = . %>%
                   filter(behavior != 'SoilMaxC'),
                lwd = 1,
                method = 'lm', color = 'black', fill = 'gray60') +
    geom_smooth(data = . %>%
                  filter(behavior == 'SoilMaxC'),
                aes(color =zone, fill = zone), 
                method = 'lm',lwd = 1 ) +
    geom_point(aes(color =zone, fill = zone, shape = zone), 
               size =2, stroke = 1, color = 'gray40'  ) +
  scale_shape_manual('Zone', values = c(21, 24)) + 
  scale_color_manual('Zone', 
                    values = c("#74A089","#E1BD6D")) +
  scale_fill_manual('Zone', 
                    values = c("#74A089","#E1BD6D")) +
  labs(x = '', y = 'Burn severity (ΔNBR)') + 
    facet_wrap(~Behavior, scales = 'free_x') +
  theme(strip.text = element_text(face = 'bold'), 
        plot.margin=margin(c(3,10,-10,3), 'em'), 
        legend.position = 'top', 
        legend.direction = 'horizontal', 
        legend.box.margin = margin(-1,0,-5,0, 'mm'))
```
  
```{r }
      bind_rows(
        BurnModels$Sev_v_MaxC2 %>% 
          as.data.frame() %>%
          rownames_to_column('Term') %>%
          tibble() %>%
          mutate(Term = replace_values(Term, 'MaxC' ~ 'Flame temperature')) %>%
          cbind(MuMIn::r.squaredGLMM(BurnModels$mc_2) %>% 
                  as_tibble() %>%
                  add_row( R2m = NA,  R2c = NA) ), 
        BurnModels$Sev_v_SoilMaxC2 %>% 
          as.data.frame() %>%
          rownames_to_column('Term') %>%
          tibble() %>%
          mutate(Term = replace_values(Term, 'SoilMaxC' ~ 'Soil Surface temperature'))%>%
          cbind(MuMIn::r.squaredGLMM(BurnModels$sc_2) %>% 
                  as_tibble() %>%
                  add_row( R2m = NA,  R2c = NA) ), 
        BurnModels$Sev_v_ROS2 %>%
        as.data.frame() %>%
          rownames_to_column('Term') %>%
          tibble() %>%
          mutate(Term = replace_values(Term, 'ros' ~ 'Rate of spread')) %>%
          cbind(MuMIn::r.squaredGLMM(BurnModels$rs_2) %>% 
                  as_tibble() %>%
          add_row( R2m = NA,  R2c = NA) ) ) %>%
        mutate(Chisq = signif(Chisq, 3), 
               Chisq = round(Chisq, 2), 
               across(c(R2m, R2c), ~ as.character(round(.x, 2)) ), 
               across(c(R2m, R2c), ~ replace_values(.x, NA ~ ''))) %>%
        rename('$\\Chi^2$' = Chisq, 
               P = `Pr(>Chisq)`, 
               `Marg. R$^2$` = R2m,
               `Cond. R$^2$` = R2c ) %>%
        mutate(P = case_when(
                  between(P,0.0009, 0.01) ~ '< 0.01',
                  P <= 0.0009 ~ '< 0.001', 
                  P >= 0.05 ~ '> 0.05', 
                  TRUE ~ as.character(round(P, 2) ) ) )  %>%
        select(-Df) %>% 
        pander::pander( )
```

# Wildfires vs. Rx burns

```{r}
BurnSeverityData %>%
  group_by(zone, type) %>%
  summarize(fires = n() ) %>%
  pivot_wider(names_from = zone, 
              values_from = fires) %>%
  mutate(Total = East + West) %>%
  pander::pander("Incidents used in this study.")
```

## Burn severity across zones and fire types

```{r WF_v_Rx, fig.width = 8, fig.height = 4, fig.cap = " "}
sev_cat <-
  tibble(Severity= c('Unburned', 'Low', 'Mod. low', 'Mod. high', 'High'), 
        dNBR = c(0.1, 0.27, 0.44, 0.66, 1.3) ) %>%
  mutate(Severity = as.factor(Severity), 
         Severity = fct_reorder(Severity, dNBR))

    BurnSeverityData  %>% 
      filter(Mean_dNBR < 0.6) %>%
      mutate(zone = fct_rev(zone), 
             zone = paste0(str_to_title(zone), 'ern zone'),
             type = fct_relevel(type, c("Wildfire",'Rx' )), 
             type = recode(type, 'Rx'= "Prescribed\nburn")) %>%
      ggplot(aes(x= type, y = Mean_dNBR)) + theme_bw(16) +
          geom_hline(data = filter(sev_cat, dNBR < 0.7),
               aes(yintercept = dNBR, 
                   lty = Severity, 
                   color = Severity) , 
               lwd = 1.1 ) + 
       geom_boxplot(varwidth = TRUE, 
                    fill = 'lightblue', 
                    staplewidth = 0.25) +
       geom_point(data = . %>%
                   group_by(type, zone) %>%
                   summarize(Mean = mean(Mean_dNBR)), 
                 aes(y = Mean), 
                 pch = 24, size = 4, 
                 color = 'white', fill = 'darkblue') + 
      labs(x = 'Fire type', 
           y = 'Burn severity (ΔNBR)') +           
      facet_wrap(~zone, scales = 'free_x') +
     coord_cartesian(ylim = c(0.08,0.65)) +
    scale_linetype_manual("Severity class", values = c(1, 3, 2, 5)) + 
    scale_color_manual("Severity class", 
                       values = c('grey80', 'grey60', 'grey40', 'grey20')) + 
    scale_y_continuous(breaks = c(0.1, 0.25, 0.44, 0.66), 
                       labels = c(0.1, 0.25, 0.44, 0.66)) + 
      theme(strip.text = element_text(face = 'bold'),
            panel.grid.major = element_blank(), 
            panel.grid.minor = element_blank(), 
            axis.text.x = element_text(color = 'black'), 
            legend.key.width = unit(2.0, 'cm'), 
            legend.key = element_rect(hjust = 0.5)) +  
    guides(linetype = guide_legend(reverse=TRUE), 
           color = guide_legend(reverse=TRUE)) 
```

```{r wf_rx_int}
  emmeans::joint_tests(BurnModels$type_comp) %>%
    tibble() %>%
    rename(`F` = F.ratio, P = p.value) %>%
    mutate(`F` = signif(F, 2),
           P = case_when(
             between(P,0.0009, 0.01) ~ '< 0.01',
             P <= 0.0009 ~ '< 0.001', 
             P >= 0.05 ~ '> 0.05', 
             TRUE ~ as.character(round(P, 2) ) ) )  %>%
  pander::pander('emmeans test indicates significant type x zone interaction.')
```

```{r wf_rx_comp}
  emmeans::joint_tests(BurnModels$type_comp, by = 'zone')  %>%
    tibble() %>%
    rename(`F` = F.ratio, P = p.value) %>%
    mutate(`F` = signif(F, 2),
           P = case_when(
             between(P,0.0009, 0.01) ~ '< 0.01',
             P <= 0.0009 ~ '< 0.001', 
             P >= 0.05 ~ '> 0.05', 
             TRUE ~ as.character(round(P, 2) ) ) )  %>%
  pander::pander('emmeans test on contrasts within the interactive linear model indicates significant difference between burn severity in wildfires vs prescribed burns in the eastern zone, but not in the western zone.')
```