---
title: Evaluating an attempt to restore summer fire in the Northern Great Plains
author: DA McGranahan & JP Angerer??
output: 
  html_document:
    highlight: tango
---

```{r setup, echo=FALSE, warning=FALSE, message = FALSE}
knitr::opts_chunk$set(message = FALSE, warning=FALSE, 
                      echo=FALSE, eval=TRUE, fig.path = '', dev='png' )
pacman::p_load(tidyverse, cowplot, patchwork, wesanderson)
  BuGnPal = c( "#67A9CF" ,"#3690C0" ,"#02818A", "#016450")
  FFgrad = wes_palette("FantasticFox1")[c(3,2,4,5)]
# load data 
  load('../data/NoFireIndices.Rdata')
  load('../data/BurnIndices.Rdata')
  load('../data/NoFireTrends.Rdata')
  load('../data/WxData.Rdata')
# some necessary things
  seasons <- 
    tibble(season = c('Spring', 'Summer'), 
           start = c(as.Date('2024-04-25'), as.Date('2024-08-07')), 
           end = c(as.Date('2024-05-25'), as.Date('2024-08-27'))) 
```

```{r}
LandsatSums <-
  NoFireTrends  %>%
    separate(ImageDate, into=c('year', 'month', 'day'), sep = '-') %>%
    mutate(season = ifelse(month=='04', 'Spring', 'Summer')) %>%
    pivot_longer(names_to = 'index', 
                 values_to = 'value', 
                 cols = c(ndvi:ndmi)) %>%
    group_by(year, season, pasture, index) %>%
    summarize(Value = mean(value), 
              .groups = 'drop') 
```

# In a nutshell

## Background 

* From 2017-2020, a project was conducted to restore spatially-patchy Rx fire in Central ND
* Two burning regimes:
  - 4 x 40-acre, full-patch burns each spring 
  - 4 x 20-acre, half-patch burns, first half in spring, second half in summer
* Mixed success in restoring fire: 
  - All spring burns were completed in each year 
  - Summer burns only completed in 2/4 years *but not for lack of trying*

## Questions

* Are there biophysical limitations on summer fires? 
  - *e.g., maybe it can just be too green and/or too wet and/or too humid to burn*
* Can trends in vegetation and weather explain why summer burns were successful in some years and not in others? 
  - *If so, maybe one can predict whether a summer burn is worth trying or not, or predicting summer wildfire risk*
* Potential implications for intentions behind growing season burns vs. pre-European summer fire regimes?
  - *Folks recognize pre-European fire regimes included considerable summer fire, and talk about greater variability in burn severity with growing-season fire as a biodiversity-friendly reason to burnin the summer. But what if pre-European summer burns were actually quite intense and burned severely because they actually only occurred under drought conditions?*

## Objectives 

* Explore patterns in fuel greenness and fire weather to determine effects on fire effectiveness
* Identify anomalies and ranges of conditions conducive to prescribed fire

## Data 

Pulled burn severity (NBR) and NDVI from Sentinel-2 imagery for 2017-2020 growing seasons.
Also pulled historical (1990-2022) NDVI from Landsat for long-term trends. 
NDVI = Normalized Difference Vegetation Index; Measure of vegetation greenness.
Data sampled from regular gridded points within burn units. 

Also retrieved historical hourly weather data, 1990-2022, from local mesonet station. 

# Results 

## Vegetation data

```{r ndvi_gg, fig.cap="Sentinel imagery summarized for the study years. A: NDVI data from four adjacent unburned pastures avoids any confounding effects of prescribed burns. X denotes seasons in which burns were not completed. B: Burn severity (dNBR) plotted against fuelbed greenness (NDVI) retrieved from the pre-burn Sentinel-2 image used for pre-fire NBR. Points represent mean for individual prescribed fires with associated standard error; circles and solid trendline indicate spring burns while broken trendline and triangles denote summer burns.", fig.height = 7, fig.width = 5}
# Get average values for burn seasons 
p1 <-
  NoFireIndices %>%
    group_by(ImageDate, pasture) %>%
    summarise(NDVI = median(ndvi) , 
              .groups = 'drop') %>%
    mutate(ImageDate = as.Date(ImageDate, '%Y-%m-%d'), 
           Year = lubridate::year(ImageDate), 
           Year = as.factor(Year), 
           SampleDate = format(ImageDate, '%m-%d'), 
           SampleDate = as.Date(SampleDate, '%m-%d'), 
           season = case_when(
             between(SampleDate, seasons$start[1], seasons$end[1]) ~ 'Spring',
             between(SampleDate, seasons$start[2], seasons$end[2]) ~ 'Summer', 
             TRUE ~ NA ) ) %>%
    filter(!is.na(season)) %>%
    group_by(season, Year) %>%
    summarize(SeasonMean = mean(NDVI), 
              SeasonSEM = sd(NDVI)/sqrt(n()), 
              .groups = 'drop') %>%
    mutate(status = case_when(
      season == 'Summer' & Year %in% c('2018', '2019') ~ "Not completed", 
      TRUE ~ "Completed") ) %>%  
    ggplot(aes(x = season, color = Year)) + theme_bw(14) +
    geom_errorbar(aes(ymin = SeasonMean - SeasonSEM, 
                      ymax = SeasonMean + SeasonSEM), 
                  width = 0.25, lwd = 0.95, 
                  alpha = 0.5,
                  position = position_dodge(width = 0.33), 
                  key_glyph = "point")  +
    geom_point(aes( y = SeasonMean, 
                    pch = interaction(status, Year)), 
               position = position_dodge(width = 0.33), 
                size = 4, stroke = 2, show.legend = F) +
    labs(x = 'Burn season', 
         y = 'NDVI', 
         ) + 
    scale_shape_manual(values = c(16, 16, 4, 16, 4, 16)) + 
   scale_color_manual(values = FFgrad) +  
    theme(axis.text.x = element_text(color = 'black')    ) +
    guides(color=guide_legend(override.aes = list(pch=15, alpha = 1, size =4))) 
p2 <-
BurnIndices  %>%
  group_by(Year, Season, fire) %>%
  summarise_at(.vars = vars(dNBR, ndvi),
               .funs = c(Mean="mean", 
                         SEM = function(x) {sd(x)/sqrt(n()) } )) %>%
  ungroup() %>%
  ggplot(aes(x = ndvi_Mean, y = dNBR_Mean, color = Year)) + theme_bw(14) +
  geom_smooth(data = . %>%
                filter(Season == 'Spring'), 
              aes(x = ndvi_Mean, y = dNBR_Mean, lty=Season), 
              method = 'lm', 
              color = 'black',
              se = T, alpha = 0.2) +
    geom_smooth(data = . %>%
                filter(Season == 'Summer'), 
              aes(x = ndvi_Mean, y = dNBR_Mean), 
              method = 'lm', 
              lty = 2,
              color = 'black',
              se = F, alpha = 0.2) +
  geom_errorbar(aes(ymin = dNBR_Mean - dNBR_SEM, 
                    ymax = dNBR_Mean + dNBR_SEM)) +
  geom_errorbarh(aes(xmin = ndvi_Mean - ndvi_SEM, 
                     xmax = ndvi_Mean + ndvi_SEM)) +
  geom_point(aes(fill = Year, 
                 shape = Season), 
             color = 'black',
             size = 3) + 
  scale_shape_manual(values=c(21, 24)) +
   scale_fill_manual(values = FFgrad) +
   scale_color_manual(values = FFgrad) +  
  labs(x = 'Pre-burn NDVI', 
       y = 'Burn severity (dNBR)') +
  theme(  legend.position = 'none', 
          legend.position.inside = c(0.8, 0.85)) 

plot_grid(p1, p2,  labels = "AUTO", 
          ncol = 1, rel_heights = c(0.8, 1))
```

```{r green_trends, fig.cap="Study season data (points) plotted within the context of 42-year historical trends (boxplots). X indicates seasons prescribed fires were not completed. A: NDVI data retrieved from Landsat imagery. B: Annual growing season rainfall accummulated by two specific dates associated with the spring and summer burn seasons, respectively.", fig.width = 5, fig.height = 7}
p3 <-
LandsatSums %>%
                            filter(year %in% c(2017:2020), 
                                   index == 'ndvi') %>%
                            group_by(year, season) %>%
                            summarize(Mean = mean(Value),
                                      SEM = sd(Value) / sqrt(n()),
                                      .groups = 'drop') %>%
                            mutate(status = case_when(
                              season == 'Summer' & 
                              year %in% c('2018', '2019') ~ "Not completed", 
                              TRUE ~ "Completed") )%>% 
  ggplot(aes(x = season)) + theme_bw(14) + 
      geom_boxplot(data = LandsatSums %>%
                     filter(index == 'ndvi')  ,
                 aes( y = Value),
                 outliers = F,
                 staplewidth = 0.25) +
  # geom_errorbar(aes( ymin = Mean - SEM,
  #                    ymax = Mean + SEM,
  #                    color = year),
  #               width = 0.5,
  #               position = position_dodge(width = 0.25)) +
         geom_point(#data = , 
        aes(y = Mean, 
                color = year, shape = status), 
             position = position_dodge(width = 0.25), 
             size = 4, stroke = 3, show.legend = T) + 
    labs(x = 'Burn season', y = 'NDVI') + 
       scale_color_manual("Year", values = FFgrad) + 
  scale_shape_manual(values = c(16, 4), guide = 'none') +

  theme(axis.text.x = element_text(color = 'black'), 
        panel.grid.major.x = element_blank() ) +
    guides(color=guide_legend(override.aes = list(pch=15, size =4)))
p4 <- 
    WxData$Rainfall %>% 
    filter(month(date) %in% c(04:09)) %>%
    separate(date, into = c('Year', 'Month', 'Day'),sep = '-') %>%
    group_by(Year) %>%
    mutate(CumPrecip = cumsum(Rainfall), 
           Month = recode(Month, '05' ="May", '08'='Aug'), 
           Month = paste0(Month, ' ', Day)) %>%
    filter(Month %in% c('May 15', 'Aug 01' )) %>%
    mutate(Month = fct_rev(Month)) %>%
    ggplot() + theme_bw(14) + 
      geom_boxplot(aes(x = Month, y = CumPrecip),
                   outliers = F,
                   staplewidth = 0.25) +
    geom_point(data = . %>%
                 filter(as.numeric(Year) %in% c(2017:2020)) %>%
                  mutate(status = case_when(
                    Month == 'Aug 01' & 
                      Year %in% c('2018', '2019') ~ "Not completed", 
                    TRUE ~ "Completed") ), 
            aes(x = Month, y = CumPrecip,
                color = Year, shape = status), 
            position = position_dodge(width = 0.25), 
            size = 4, stroke = 3, show.legend = T) + 
  labs(x = 'Date', y = 'Cumulative Rainfall (mm)') + 
  scale_shape_manual(values = c(16, 4), guide = 'none') +
  scale_color_manual("Year", values = FFgrad) + 
  theme(axis.text.x = element_text(color = 'black'), 
        panel.grid.major.x = element_blank() ) +
  guides(color=guide_legend(override.aes = list(pch=15, size =4)))

p3 / p4 +
  patchwork::plot_layout(
    guides = "collect",
    nrow = 2
  ) +
  patchwork::plot_annotation(
    tag_levels = "A"
  ) &
  theme(legend.position = "right", legend.direction = "vertical")
```

## Weather data 

```{r weather, fig.cap="Four fire weather variables plotted by burn day values (open triangles), study season means (open circles and X for seasons without completed burns), and 42-year historical trends (boxplots).", fig.height = 9}
RainyDays <- 
   WxData$Rainfall %>%
      filter( ! is.na(season), Rainfall > 0.1) 
# Seasonal trends over time   
  WxData$HourlyWx %>%
    filter(! date %in% RainyDays$date, # remove rainy days with no chance of burning
           ! is.na(season)) %>%
    group_by(Year, date, season) %>%
    summarise_at(.vars = vars(c(DewPoint, RelHum, WindSpeed,VPD)),
                 .funs = c("mean")) %>%
    ungroup() %>%
    pivot_longer(names_to = 'variable', 
                 values_to = 'value', 
                 cols = DewPoint:VPD) %>%
    group_by(Year, season, variable) %>%
    summarise(Value = mean(value), 
              .groups = 'drop') %>%
    mutate(variable = recode(variable, 
                             'DewPoint' = 'Dew Point (°C)', 
                             'RelHum' = 'Relative Humidity (%)', 
                             'VPD' = 'Vapor Pressure Deficit', 
                             'WindSpeed' = 'Wind speed (m/s)') , 
           Year = as.character(Year), 
           season = fct_rev(season)) %>%
    ggplot() + theme_bw(16) + 
      geom_boxplot( aes(x=season, 
                       y = Value),
                   outliers = F, size = 0.75,
                   staplewidth = 0.25) +
      geom_point(data = . %>%
                   filter(Year %in% c(2017:2020)) %>%
                   mutate(status = case_when(
                     season == 'Summer' & Year %in% c('2018', '2019') ~ "Not completed", 
                     TRUE ~ "Completed") ) , 
                 aes(x = season, y = Value, 
                     color = Year, shape = status), 
                 position = position_nudge(x = -.2), 
                  fill = 'grey90', 
                 size = 3, stroke = 2, show.legend = T) + 
    geom_point(data = WxData$BurnDayWx , 
               aes(x = season, y = value, color =Year), 
               position = position_nudge(x = +.2), 
               shape = 24, fill = 'grey90', 
               size = 3, stroke = 2,
                show.legend = F) + 
      labs(x = 'Burn season', y = 'Weather variable value') + 
      facet_wrap(~variable, scales = 'free_x', ncol = 1)  +
      scale_shape_manual(values = c(21, 4), guide = 'none') +
    scale_color_manual(values = FFgrad) +  
      theme(axis.text.y = element_text(color = 'black', angle = 45, hjust = 0.5), 
            panel.grid.major.y = element_blank() , 
            strip.text = element_text(face='bold'), 
            strip.background = element_rect(fill = '#A1D99B')) +
    coord_flip() + 
      guides(color=guide_legend(override.aes = list(pch=15, size =4)))
```

# Supplemenary Info

## Seasonal trends

Summer fires were successful in 2017 and 2020. 
Vegetation in both years was less green and rainfall lower.

```{r ndvi_trend, eval = TRUE, fig.cap="Green shaded areas span the first and last dates of completed burns in the spring and summer.", fig.height = 4, fig.width = 8}
NoFireIndices %>%
  group_by(ImageDate, pasture) %>%
    summarise(Mean = mean(ndvi), 
              SEM = sd(ndvi)/sqrt(n()), 
              .groups = 'drop') %>%
  mutate(ImageDate = as.Date(ImageDate, '%Y-%m-%d'), 
         Year = lubridate::year(ImageDate), 
         Year = as.factor(Year), 
         SampleDate = format(ImageDate, '%m-%d'), 
         SampleDate = as.Date(SampleDate, '%m-%d') ) %>%
  filter(Year != 2016) %>%
ggplot(aes(x = SampleDate, 
           color = Year, 
           fill = Year, 
           group = interaction(pasture, Year))) + theme_bw(14) +
  geom_rect(data= seasons, inherit.aes = FALSE,
              aes(xmin=start, xmax=end, ymin=-Inf, ymax=+Inf), 
              fill='#A1D99B', alpha=0.5) +
  geom_smooth(aes(y = Mean, group = Year),
              se = T,
              level = 0.99) +
  geom_errorbar(aes(ymin = Mean - SEM, 
                    ymax = Mean + SEM),
                alpha = 0.5) +
  geom_line(aes(y = Mean),
             alpha = 0.5) +
  labs(x = 'Image date', y = 'NDVI') + 
    scale_fill_manual(values = FFgrad) + 
  scale_color_manual(values = FFgrad) + 
  scale_x_date( date_breaks = 'month', 
                date_labels = '%b-%d', 
                minor_breaks = NULL) 
```

```{r precip_trend, eval = TRUE, fig.cap="Thick black line shows smoothed 40-yr trend. Green shaded areas span the first and last dates of completed burns in the spring and summer.", fig.height = 4, fig.width = 8}
  WxData$Rainfall %>% 
    filter(month(date) %in% c(04:09)) %>%
    separate(date, into = c('Year', 'Month', 'Day'),sep = '-') %>%
    group_by(Year) %>%
    mutate(CumPrecip = cumsum(Rainfall)) %>%
    ggplot(aes(x = day)) + theme_bw(14) +
    geom_rect(data= seasons, inherit.aes = FALSE,
              aes(xmin=start, xmax=end, ymin=-Inf, ymax=+Inf), 
              fill='#A1D99B', alpha=0.5) +
    geom_line(aes(y = CumPrecip, group = Year),
              color = 'grey40', 
              alpha = 0.5) +
    geom_smooth(aes(y = CumPrecip),
                se = F,
                color = 'black',
                size = 1,
                level = 0.99) +
    geom_line(data = . %>%
                filter(as.numeric(Year) %in% c(2017:2020)) , 
              aes(y = CumPrecip, 
                  color = Year,
                  group = Year),
              size = 2) +
    labs(x = 'Day of year', y = 'Cumulative rainfall (mm)') + 
    scale_fill_manual(values = FFgrad) + 
    scale_color_manual(values = FFgrad) + 
    scale_x_date( date_breaks = 'month', 
                  date_labels = '%b-%d', 
                  minor_breaks = NULL)
```

```{r veg_trends, eval = FALSE, fig.cap="Blue shaded areas span the first and last dates of completed burns in the spring and summer.", fig.height = 8, fig.width = 10}
NoFireIndices %>%
  pivot_longer(names_to = 'index', 
               values_to = 'value', 
               cols = c(ndvi:ndmi)) %>%
  group_by(ImageDate, pasture, index) %>%
    summarise(Mean = mean(value), 
              SEM = sd(value)/sqrt(n()), 
              .groups = 'drop') %>%
  mutate(ImageDate = as.Date(ImageDate, '%Y-%m-%d'), 
         Year = lubridate::year(ImageDate), 
         Year = as.factor(Year), 
         SampleDate = format(ImageDate, '%m-%d'), 
         SampleDate = as.Date(SampleDate, '%m-%d'), 
         index = toupper(index), 
          index = fct_rev(index)) %>%
  filter(Year != 2016) %>%
ggplot(aes(x = SampleDate, 
           color = Year, 
           fill = Year, 
           group = interaction(pasture, Year))) + theme_bw(14) +
  geom_rect(data= seasons, inherit.aes = FALSE,
              aes(xmin=start, xmax=end, ymin=-Inf, ymax=+Inf), 
              fill='#A1D99B', alpha=0.5) +
  geom_smooth(aes(y = Mean, group = Year),
              se = T,
              level = 0.99) +
  geom_errorbar(aes(ymin = Mean - SEM, 
                    ymax = Mean + SEM),
                alpha = 0.5) +
  geom_line(aes(y = Mean),
             alpha = 0.5) +
  labs(x = 'Image date', y = 'Index value') + 
  facet_wrap(~index, ncol=1, scale = 'free_y') +
    scale_fill_manual(values = FFgrad) + 
  scale_color_manual(values = FFgrad) + 
  scale_x_date( date_breaks = 'month', 
                date_labels = '%b-%d', 
                minor_breaks = NULL) 
```

```{r veg_by_season, eval = FALSE, fig.cap="Data summarized by year from shaded blue ranges above.", fig.height = 4, fig.width = 10}
# Get average values for burn seasons 
  NoFireIndices %>%
    pivot_longer(names_to = 'index', 
                 values_to = 'value', 
                 cols = c(ndvi:ndmi)) %>%
    group_by(ImageDate, pasture, index) %>%
    summarise(Median = median(value) , 
              .groups = 'drop') %>%
    mutate(ImageDate = as.Date(ImageDate, '%Y-%m-%d'), 
           Year = lubridate::year(ImageDate), 
           Year = as.factor(Year), 
           SampleDate = format(ImageDate, '%m-%d'), 
           SampleDate = as.Date(SampleDate, '%m-%d'), 
           season = case_when(
             between(SampleDate, seasons$start[1], seasons$end[1]) ~ 'Spring',
             between(SampleDate, seasons$start[2], seasons$end[2]) ~ 'Summer', 
             TRUE ~ NA ) ) %>%
    filter(!is.na(season)) %>%
    group_by(season, Year, index) %>%
    summarize(SeasonMean = mean(Median), 
              SeasonSEM = sd(Median)/sqrt(n()), 
              .groups = 'drop') %>%
    mutate(status = case_when(
      season == 'Summer' & Year %in% c('2018', '2019') ~ "Not completed", 
      TRUE ~ "Completed"), 
          index = toupper(index), 
          index = fct_rev(index)) %>%
    ggplot(aes(x = season, y = SeasonMean, color = Year)) + theme_bw(14) +
    geom_errorbar(aes(ymin = SeasonMean - SeasonSEM, 
                      ymax = SeasonMean + SeasonSEM), 
                  width = 0.25, lwd = 0.95, 
                  position = position_dodge(width = 0.33), 
                  key_glyph = "point")  +
    geom_point(aes(pch = interaction(status, Year)), 
               position = position_dodge(width = 0.33), 
                size = 4, stroke = 2, show.legend = F) +
    labs(x = 'Burn season', 
         y = 'Index value', 
         caption = 'X = Year/season in which burns were not completed') + 
    scale_shape_manual(values = c(16, 16, 4, 16, 4, 16)) + 
    facet_wrap(~index, ncol = 3, scale = "free") +
    theme(axis.text.x = element_text(color = 'black')) +
    guides(color=guide_legend(override.aes = list(pch=15, size =4))) 
```

```{r sev_veg, eval = FALSE, fig.cap="Vegetation indices pulled from closest pre-burn image, same as pre-burn NBR.", fig.height = 3.8, fig.width = 10}
BurnIndices  %>%
  pivot_longer(names_to = 'name', 
               values_to = 'index', 
               cols = c(ndvi, msi, ndmi)) %>%
  group_by(Year, Season, fire, name) %>%
  summarise_at(.vars = vars(dNBR, index),
               .funs = c(Mean="mean", 
                         SEM = function(x) {sd(x)/sqrt(n()) } )) %>%
  ungroup() %>%
  mutate(name = toupper(name), 
         name = fct_rev(name)) %>% 
  filter(Season == "Spring") %>%
  ggplot(aes(x = index_Mean, y = dNBR_Mean, color = Year)) + theme_bw(14) +
  geom_smooth(aes(x = index_Mean, y = dNBR_Mean), 
              method = 'lm', 
              color = 'black',
              se = T, alpha = 0.2) +
  geom_errorbar(aes(ymin = dNBR_Mean - dNBR_SEM, 
                    ymax = dNBR_Mean + dNBR_SEM)) +
  geom_errorbarh(aes(xmin = index_Mean - index_SEM, 
                     xmax = index_Mean + index_SEM)) +
  geom_point() + 
  labs(x = 'Pre-burn index value', 
       y = 'Burn severity (dNBR)') + 
  geom_smooth(method = 'lm', se = F) +
  facet_wrap(~name, scales = 'free')
```

```{r, eval = FALSE}
full_join(by = 'fire', 
fires %>%
  as_tibble() %>%
  filter(Year == '2018') %>%
  select(fire, PreBurn), 

BurnIndices  %>%
  pivot_longer(names_to = 'name', 
               values_to = 'index', 
               cols = c(ndvi, msi, ndmi)) %>%
  group_by(Year, Season, fire, name) %>%
  summarise_at(.vars = vars(dNBR, index),
               .funs = c(Mean="mean", 
                         SEM = function(x) {sd(x)/sqrt(n()) } )) %>%
  ungroup() %>%
  filter(Year == '2018') %>%
  select(fire, name, dNBR_Mean, index_Mean) %>%
  pivot_wider(names_from = name, 
              values_from = index_Mean) ) %>%
  rename(dNBR = dNBR_Mean) %>%
  arrange(desc(msi)) 
```


## Historical trends 

```{r eval = FALSE}
LandsatSums %>%
  filter(year %in% c(2017:2020)) %>%
  group_by(year, season, index) %>%
  summarize(Mean = mean(Value),
            SEM = sd(Value)/sqrt(n()), 
            .groups = 'drop') %>%
  mutate(status = case_when(
    season == 'Summer' & year %in% c('2018', '2019') ~ "Not completed", 
    TRUE ~ "Completed"), 
    index = toupper(index), 
    index = fct_rev(index) ) %>%
    ggplot(aes(x=season)) + theme_bw(14) + 
    geom_boxplot(data = LandsatSums %>% mutate(index = toupper(index), 
                                               index = fct_rev(index)),
                         aes( y = Value),
                         outliers = F, 
                 staplewidth = 0.25) +
    geom_point(aes(y = Mean, color = year, shape = status), 
               position = position_dodge(width = 0.25), 
               size = 3, stroke = 2, show.legend = F) + 
    facet_wrap(~index, scales = 'free_y')  +
  scale_shape_manual(values = c(16, 4)) +
  theme(axis.text.x = element_text(color = 'black'), 
        panel.grid.major.x = element_blank() )
```

```{r }
StudySums <- 
  LandsatSums %>%
    filter(year %in% c(2017:2020)) %>%
    group_by(year, season, index) %>%
    summarize(Mean = mean(Value),
              SEM = sd(Value)/sqrt(n()), 
              .groups = 'drop') %>%
    mutate(status = case_when(
      season == 'Summer' & year %in% c('2018', '2019') ~ "Not completed", 
      TRUE ~ "Completed"), 
      index = toupper(index), 
      index = fct_rev(index) ) 

HistoricalSpectralData <- 
  bind_rows(
  LandsatSums %>%
  filter(year %in% c(1990:1999))%>% 
  mutate(range = '1990-1999', 
         index = toupper(index), 
         index = fct_rev(index)), 
  LandsatSums %>%
    filter(year %in% c(2012:2022))%>% 
    mutate(range = '2012-2022', 
           index = toupper(index), 
           index = fct_rev(index)), 
  LandsatSums %>%
    mutate(range = '1990-2022', 
           index = toupper(index), 
           index = fct_rev(index)) )
```

```{r historical_ndvi, eval = FALSE, fig.width = 7, fig.height =3}
  ggplot() + theme_bw(14) + 
  geom_boxplot(data = HistoricalSpectralData  %>% filter(index == 'NDVI'),
               aes(x=season, 
                   y = Value),
               outliers = F, 
               staplewidth = 0.25) +
  # geom_errorbar(aes( ymin = Mean - SEM, 
  #                    ymax = Mean + SEM, 
  #                    color = year), 
  #               width = 0.5, 
  #               position = position_dodge(width = 0.25)) + 
  geom_point(data = StudySums %>% filter(index == 'NDVI'), 
            aes(x = season, y = Mean, 
                color = year, shape = status), 
             position = position_dodge(width = 0.25), 
             size = 3, stroke = 2, show.legend = T) + 
    labs(x = 'Burn season', y = 'NDVI') + 
  facet_wrap(~range)  +
  scale_shape_manual(values = c(16, 4), guide = 'none') +
    scale_color_manual(values = FFgrad) +
  theme(axis.text.x = element_text(color = 'black'), 
        panel.grid.major.x = element_blank() ) +
    guides(color=guide_legend(override.aes = list(pch=15, size =4)))
```

```{r historical_phenology, eval = TRUE, fig.width = 8, fig.height =11.5}
WxData$HourlyWx %>%
  filter(! date %in% RainyDays$date, # remove rainy days with no chance of burning
         ! is.na(season)) %>%
  group_by(Year, date, season) %>%
  summarise_at(.vars = vars(c(DewPoint, RelHum, WindSpeed,VPD)),
               .funs = c("mean")) %>%
  ungroup() %>%
  pivot_longer(names_to = 'variable', 
               values_to = 'value', 
               cols = DewPoint:VPD) %>%
  group_by(Year, season, variable) %>%
  summarise(Value = mean(value), 
            .groups = 'drop') %>%
  mutate(Year = as.character(Year), 
         #season = fct_rev(season), 
         period = case_when(
           Year %in% c(1990:1999) ~  '1990-1999', 
           Year %in% c(2012:2022) ~  '2012-2022', 
           TRUE ~ NA
         )) %>%
  filter(!is.na(period)) %>%
  bind_rows(WxData$Rainfall %>% 
              filter(month(date) %in% c(04:09)) %>%
              separate(date, into = c('Year', 'Month', 'Day'),sep = '-') %>%
              group_by(Year) %>%
              mutate(CumPrecip = cumsum(Rainfall), 
                     Month = recode(Month, '05' ="May", '08'='Aug'), 
                     season = paste0(Month, ' ', Day)) %>%
              ungroup() %>%
              filter(season %in% c('May 15', 'Aug 01' )) %>%
              mutate(season = ifelse(season == 'May 15', 'Spring', 'Summer'), 
                     period = case_when(
                       Year %in% c(1990:1999) ~  '1990-1999', 
                       Year %in% c(2012:2022) ~  '2012-2022', 
                       TRUE ~ NA
                     ) )  %>%
              filter(!is.na(period)) %>%
              select(Year, season, CumPrecip, period) %>%
              pivot_longer(names_to = 'variable', 
                           values_to = 'Value', 
                           cols = CumPrecip))  %>%
  bind_rows(HistoricalSpectralData  %>% 
              select(-pasture) %>%
              filter(index == 'NDVI', 
                     range != '1990-2022') %>%
              rename(Year = year, period = range, variable = index)
            ) %>%
  mutate( variable = factor(variable, levels = c('NDVI', 'CumPrecip', 'DewPoint', 
                                                 'RelHum', 'VPD', 'WindSpeed')), 
          variable = recode(variable, 
                           'NDVI' = "Fuelbed\ngreenness\n(NDVI)", 
                           'CumPrecip' = 'Cumulative\nrainfall\n(mm)',
                           'DewPoint' = 'Dew\nPoint\n(°C)', 
                           'RelHum' = 'Relative\nHumidity\n(%)', 
                           'VPD' = 'Vapor\nPressure\nDeficit', 
                           'WindSpeed' = 'Wind\nspeed\n(m/s)'), 
          period = recode(period, '1990-1999' = '1990-99', '2012-2022' = '2012-22')) %>%
  ggplot() + theme_bw(16) + 
  geom_hline(data = . %>%
               filter(as.numeric(Year) %in% c(2017:2020)) %>%
               mutate(status = case_when(
                 season == 'Summer' & 
                   Year %in% c('2018', '2019') ~ "Not completed", 
                 TRUE ~ "Completed") ) %>%
               group_by(variable, Year, season, status) %>%
               slice(1) , 
             aes(yintercept = Value,
                 color = Year, lty = status), 
             lwd = 1 ) + 
  geom_boxplot(aes(x=period, 
                   y = Value),
               outliers = F, 
               staplewidth = 0.25) +
  labs(x = 'Historical period', y = ' ') + 
  facet_grid(variable~season, scales = "free_y")  +
  scale_linetype_manual("Burn status", values = c(1,2)) +
  scale_color_manual("Study year", values = FFgrad) +
  theme(axis.text.x = element_text(color = 'black'), 
        panel.grid.major.x = element_blank() , 
        strip.text.y = element_text(angle = 0), 
        strip.background = element_rect(fill = '#A1D99B')) 
```

```{r eval = FALSE}
HistoricalSpectralData  %>% 
  filter(index == 'NDVI', 
         ! is.na(Value), 
         range != '1990-2022') %>%
  group_by(season, range) %>%
  summarise(boxplot= list(setNames(boxplot.stats(Value)$stats,                                      c('bottom_whisker','bottom_hinge','median','top_hinge','top_whisker'))), 
              .groups = 'drop') %>%
    unnest_wider(boxplot) %>%
  mutate(across(c(bottom_whisker:top_whisker), ~ round(., 2)) ) %>%
  pander::pander()
  #   summarise(across(Value, 
  #                    list(min=min, Q1=~quantile(., probs = 0.25),
  #                         median=median, Q3=~quantile(., probs = 0.75),
  #                         max=max),  
  #                    .names = "{.fn}"), 
  #             .groups = 'drop') %>%
  # mutate(across(c(min:max), ~ round(., 2)) ) %>%
  # pander::pander()
```
